Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-08 prop_letter.tex
prop_letter.tex \section{Born} Updating...

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state. When we shift our attention to QIVPMs, we also want to construct
QIVPMs according to a simple rule. Since the ranges of the quantum
probability measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$,
respectively, if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to an interval~$\iota\left(x\right)\in\mathscr{I}$, then we can
concatenate it with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a $\mathscr{I}$-valued function~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although $\bar{\mu}$ and QIVPMs have the same the domain and range,
$\bar{\mu}$ is not necessary a QIVPM. For example, the following
theorem highlight a situation where $\bar{\mu}$ cannot be a QIVPM.

@@

Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\textrm{ .}
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{equation}
\begin{aligned} & \mathrel{\hphantom{\nsubseteq}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\textrm{ ,}
\end{aligned}
\end{equation}
and cannot be a QIVPM.

In other words, even if $\iota\circ\mu$ has the same domain and range
as a QIVPM, we may need to verify whether $\iota\circ\mu$ is a QIVPM
individually. For example, in the four-dimensional Hilbert space,
\begin{equation}
\iota\left(x\right)=\begin{cases}
\imposs & \textrm{, if }x=0\textrm{ ;}\\
\left[0,\tfrac{1}{2}\right] & \textrm{, if }x\in\left(0,\tfrac{1}{4}\right]\textrm{ ;}\\
\unknown & \textrm{, if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\textrm{ ;}\\
\left[\tfrac{1}{2},1\right] & \textrm{, if }x\in\left[\tfrac{3}{4},1\right)\textrm{ ;}\\
\necess & \textrm{, if }x=1\textrm{ .}
\end{cases}
\end{equation}
and $\rho=\left(\proj{0}+\proj{1}+\proj{2}+\proj{3}\right)/4$ inducing
a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ is a non-trivial
fact~\citep{HOSTunpublished}. In order to verify whether $\iota\circ\mu$
is a QIVPM easier, we can impose some conditions on $\iota$. Notice
that our counterexamples are in a four-dimensional Hilbert space,
the conditions for $\iota$ for a Hilbert space of dimension~$d\le3$
is not as strict as the general conditions.

\begin{thm}In a Hilbert space of dimension~$d\le3$, if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfying for any three numbers~$x_{0}$
and $x_{1}\in\left[0,1\right]$, $x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\textrm{ .}\label{eq:iota-add}
\end{equation}
Then, $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always a QIVPM
for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.\end{thm}

>>>>

Recall in Sec.~\ref{sec:fuzzy}, we illustrate the non-constructive
postulates of the quantum probability measures first, and then introduce
the Born rule which induces a quantum probability measures from each
state. When we shift our attention to QIVPMs, we also want to construct
QIVPMs according to a simple rule. Since the ranges of quantum probability
measures and QIVPMs are $\left[0,1\right]$ and $\mathscr{I}$, respectively,
if there is a function~$\iota$ mapping a number~$x\in\left[0,1\right]$
to an interval~$\iota\left(x\right)\in\mathscr{I}$, then we can
concatenate it with a quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
and get a $\mathscr{I}$-valued function~$\bar{\mu}:\events\overset{\mu}{\rightarrow}\left[0,1\right]\overset{\iota}{\rightarrow}\mathscr{I}$.
Although $\bar{\mu}$ and QIVPMs have the same domain and range, $\bar{\mu}$
is not necessary a QIVPM. For example, the following theorem highlight
a situation where $\bar{\mu}$ cannot be a QIVPM.

@@

Consider the state
\begin{equation}
\begin{aligned}\rho={} & x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}\\
 & +\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}\,.
\end{aligned}
\end{equation}
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{equation}
\begin{aligned} & \mathrel{\hphantom{\nsubseteq}}\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{aligned}
\end{equation}
and cannot be a QIVPM.

In other words, even if $\iota\circ\mu$ has the same domain and range
as a QIVPM, we may need to verify whether $\iota\circ\mu$ is a QIVPM
individually. For example, in the four-dimensional Hilbert space,
\begin{equation}
\iota\left(x\right)=\begin{cases}
\imposs\,, & \textrm{ if }x=0\,;\\
\left[0,\tfrac{1}{2}\right]\,, & \textrm{ if }x\in\left(0,\tfrac{1}{4}\right]\,;\\
\unknown\,, & \textrm{ if }x\in\left(\tfrac{1}{4},\tfrac{3}{4}\right)\,;\\
\left[\tfrac{1}{2},1\right]\,, & \textrm{ if }x\in\left[\tfrac{3}{4},1\right)\,;\\
\necess\,, & \textrm{ if }x=1\,.
\end{cases}
\end{equation}
and $\rho=\left(\proj{0}+\proj{1}+\proj{2}+\proj{3}\right)/4$ inducing
a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$ is a non-trivial
fact~\citep{HOSTunpublished}. In order to verify whether $\iota\circ\mu$
is a QIVPM easier, we can impose some conditions on $\iota$. Notice
that our counterexamples are in a four-dimensional Hilbert space,
the conditions for $\iota$ for a Hilbert space of dimension~$d\le3$
is not as strict as the general conditions.

\begin{thm}In a Hilbert space of dimension~$d\le3$, if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$
subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfying for any three numbers~$x_{0}$
and $x_{1}\in\left[0,1\right]$, $x_{0}+x_{1}\in\left[0,1\right]$
implies
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
Then, $\iota\circ\mu:\events\rightarrow\mathscr{I}$ is always a QIVPM
for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.\end{thm}
